### 2015-04-16 rapport/formalism.tex
rapportv1_160415

%-------------------------------------
\subsection{Parallelization}
\label{sect:parall}
% Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.

% Maybe the more important parallelization technique of such simulations is called \textit{SPMD} (Single Program, Multiple Data) which can also be called \textit{coarse-grain data parallelism}. This concept represents the well-known and broadly-used domain decomposition technique to split the mesh in balanced sub-meshes for available resources. Because of the good locality degree of numerical simulations, only a small number of additional elements, called a \textit{recorvery} or a \textit{ghost} area, are needed to exchange information computed by another resource to correctly compute stencils. As the amount of communications of such parallelization are rather small compared to the amount of computations (small neighborhood), and also by the addition of recovery of communications by computations, this parallelization technique has proved many times its efficiency and scalability on numerical simulations~\cite{}. This parallelization technique can be applied on shared and distributed memory architectures (clusters, multi-core CPUs), and it stays the most efficient parallelization technique if the numerical simulation has to be executed on a distributed memory architecture. For this reason, this parallelization stays a reference in the domain and is still studied for more complex simulations, using unstructured meshes, multi-grids techniques etc.

>>>>

%-------------------------------------
\subsection{Parallelization}
\label{sect:parall}
Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.

The \textit{SPMD} (Single Program, Multiple Data) parallelization technique which is called \textit{coarse-grain data parallelism} in the rest of this paper, is broadly used because of its scalability. This parallelization concept is the well-known domain decomposition technique where the mesh is splitted in balanced sub-meshes for available resources, and the same program is applied on each sub-part of the initial mesh. Because of the good locality degree of numerical simulations, only a small number of additional elements, called a \textit{recorvery} or a \textit{ghost} area, are needed to get information computed on another resource. This ghost area is however unavoidable to correctly compute stencils because of the neighborhood $\mathcal{N}$. As the amount of data to synchronize between resources is rather small compared to the amount of computation (small neighborhood), and also by the addition of recovery of synchronizations by computations, this parallelization technique has proved many times its efficiency and scalability on numerical simulations~\cite{}. This parallelization technique can be applied on shared and distributed memory architectures (clusters, multi-core CPUs), and it is one of the most efficient parallelization technique if the numerical simulation has to be executed on a distributed memory architecture such as a cluster. Figure~\ref{} illustrates this parallelization technique.

More formally, in a SPMD parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $n$ balanced sub-meshes (for $n$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{n-1}\}$. This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper. The second step of the parallelization is to identify in the ordered list of computation $\Gamma$ when synchronizations are needed between resources.

\medskip
\begin{mydef}
\textit{A \textit{synchronisation} is needed between two computations $c_1$ and $c_2$, denoted $c_1 \ll c_2$, if $c_2=s_2(R_2,w_2,D_2,e_2)$, and $w_1 \cap R_2 \neq \emptyset$.}
\end{mydef}

Actually, a synchronization can only be needed by the data read in a stencil computation, and only if this data has been modified before, which means that it has been written before.
A new ordered list of computations $\Gamma'$ is obtained from the insertion of the synchronizations at the good places. The final step of this parallelization is to run the same ordered list of computation $\Gamma'$ on each resource. For each resource $0 \leq k \leq n-1$ a multi-stencil program defined by
\begin{equation}
\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma'),
\end{equation}
is runned.

PARALLEL,SYNC definitions + algorithm
